\documentclass[a4paper,12pt]{article}
\usepackage{amsmath}
\usepackage{enumerate}

\pagestyle{empty} \setlength{\parindent}{0mm}
\addtolength{\topmargin}{-0.5in} \setlength{\textheight}{9in}
\addtolength{\textwidth}{1in} \addtolength{\oddsidemargin}{-0.5in}


\begin{document}
\title{Assignment 4}
\author{Matt Forbes}
\date{February 14, 2011}
\maketitle

\section*{What's Working}
\begin{enumerate}[]
  \item My server passes all the tests executed by the test scripts,
    so that's a good sign. Security is enforced by matching the
    hostname of the client to a line in the .rfahosts file. At the
    initial run of the program, a linked list of strings is made that
    can be quickly iterated over to find a match. IP addresses are
    only used in debugging messages, only the hostname is
    used. Hopefully that is the correct behavior.
    
  \item Opening files also seems to be working as expected. I haven't
    thoroughly tested that the mode passed to open is correct, but
    it's worked for everything I've tried. All the flags are being
    OR'd appropriately, and exactly one of 'R' 'W' and 'RW' must be
    present or an error is returned. Also, file names with spaces are
    parsed correctly. The simple validation of pathnames is also being
    performed.

  \item Reading is pretty straightforward, and has worked well as long
    as open is functioning.

  \item Write also works appropriately.
    
  \item Close works.
\end{enumerate}

\section*{What's Not Working}
\begin{enumerate}[]
  \item Nothing seems broken, although there are definitely
    improvemnts to be made. Initially, I started using strtok() to
    parse the request strings, but after realizing it wouldn't work
    for the Open command without some extra work I reverted to simple
    char-by-char processing. While this is working for now, it's not
    quite robust enough and could be prone to buffer overflows if
    name_length is specified incorrectly. 
    
  \item Also, I'm not 100\% sure that all the correct error numbers
    are being returned. I've been counting on the global errno being
    set appropriately in most cases, which is hopefully enough.
\end{enumerate}

\end{document}
