\documentclass[a4paper,12pt]{article}
\usepackage{amsmath}
\usepackage{enumerate}

\pagestyle{empty} \setlength{\parindent}{0mm}
\addtolength{\topmargin}{-0.5in} \setlength{\textheight}{9in}
\addtolength{\textwidth}{1in} \addtolength{\oddsidemargin}{-0.5in}


\begin{document}
\title{245 - Homework 6}
\author{Matt Forbes}
\date{December 4, 2010}
\maketitle

\pagebreak

\section*{Timesheet}

\begin{tabular}{r l}

  Nov 14 & 3:30pm to 1:00am\\

  Nov 20 & 5pm to 1:30am \\

  Nov 21 & 9:30am to 11:30am \\

  Nov 22 & 7:00pm to 2:00am \\

  Nov 23 & 7:00pm to 11:00pm \\

  Nov 24 & 9:00pm to 11:00pm \\

  Nov 25 & 9:00am to 12:00pm and 9:00pm to 11:00pm \\

  Nov 26 & 9:00am to 1:00pm and 3:00pm to 5:00pm and 8:00pm to 10:00pm \\

  Nov 27 & 11:00am to 2:00pm \\

  Nov 30 & 7:00pm to 12:00am
  
\end{tabular}

\section*{Summary}

All features outlined in the game specification are implemented.

\begin{enumerate} [a)]

  \item Level construction.

  \item Player's hunger is recorded and is responded to at the various
    levels. Including healing every 10 steps at positive hunger and
    subtracting one health every 5 steps at negative hunger.

  \item Both drinks of health/sickness are available in the
    game. Before drinking a drink, any drinks found are reported as
    unknown. Upon drinking one of any kind, the type of drink is
    revealed for the rest of the game. The effects of the drinks are
    implemented as wel.

  \item Traps are invisible until stepped on, and can be activated
    again when visible. Both types of traps work as specified.

  \item Movement order is player then monsters.

  \item Player can obtain gold by walking on to a space which has gold
    on it. The amount of gold is a random (between 5 and 100.)

  \item Monsters' health is determined by which level they are (and a
    random factor). They can heal themselves by one health point for
    every 15 steps the player makes (regardless of hunger/health/etc.)

  \item The movement of monsters is different depending on whether the
    player is in the same room. The monster has a high chance of
    moving towards the player if he is in the room, and is completely
    random if he is not.

  \item Combat occurs when either the player moves to a space occupied
    by a monster or the reverse. The attacker has the upper hand but
    it is still randomly decided (based on specification.) There can
    only be one combat event between the player and a monster per
    turn, but multiple monsters can attack the player in the same
    turn.

  \item The player's inventory can hold up to 10 items. He can drop
    and item anywhere he is standing. Only food can be eaten, and
    drinks can be drunk.

  \item Visibility rules are defined based on line of sight. These
    rules are just how the specification describes them and how they
    work in the example game.

  \item The game is over when the player's health drops below zero or
    he exits through the up-stairs on the first level. If he leaves
    with the diamond in his inventory he wins, any other termination
    case means he lost.

  \item All game rules are implemented even if not specifically
    outlined here!
    
\end{enumerate}

\end{document}
