\documentclass[a4paper,12pt]{article}
\usepackage{amsmath}
\usepackage{enumerate}

\pagestyle{empty} \setlength{\parindent}{0mm}
\addtolength{\topmargin}{-0.5in} \setlength{\textheight}{9in}
\addtolength{\textwidth}{1in} \addtolength{\oddsidemargin}{-0.5in}


\begin{document}
\title{Homework 4}
\author{Matt Forbes}
\date{November 16, 2010}
\maketitle

\section*{Problem 1}


\begin{enumerate}[a)]
  \item

    A linked list provides a very good representation of the mechanics
    of the game. The list is the 'circle' that the kids are lined up
    in, and simulating the game is as easy as iterating through the
    list and removing every $w^{th}$ kid.

    
    To start with, we would need to build the list. Each kid gets his
    own node in the list, and they are inserted in order (where order
    is determined by the circle, original caller is 1, his right is 2,
    etc.). This would take $O(p)$ time as there are p kids.

    
    Simulating the selection requires $p-1$ deletions from the
    list. One for each kid that is eliminated. Starting from the head
    of the list, follow $w-1$ links in the list. Then delete the next
    link. Now, follow $w-1$ links again, and then delete the
    next. Whenever you reach the end of the list, wrap around to the
    first element (counting that as one visit). Continue doing this
    until there is only one link left in the list. That last element
    remaining is the kid who is ``it.'' In this step, we followed
    $w-1$ pointers exactly $p-1$ times, which is $O(pw)$.

    
    Adding the running times of both the creation of the list and
    elimination phase gives $O(p + pw)$ which is more simply just
    $O(pw)$. 

  \item Using OS\_Tree developed in class to implement emmm\_select:
    
\subsection*{Pseduo Code}

    \begin{verbatim}
      (* emmm_select performs eeny-meeny-miny-mo selection on the 
      array of p kids (kid_arr), using a song with w words *)

      def emmm_select(kids, p, w):
          T is a new OS_Tree of size p
          for i in 1...p:
              insert(T, kids[i])

          caller = 1
          size = p
          while size > 1:
              (* kid going out is w-1 kids away from the caller, with wrapping *)
              out = (caller + w-1) % (total + 1)
              size = size - 1
              kid = OS_Select(T, out)
              remove(T, kid)

              (* next caller is kid who takes out's position in circle *)
              caller = out

          (* return rank 1 (only) kid left in tree *)
          return OS_Select(T, 1)

    \end{verbatim}

\subsection*{Justification}

emmm\_select relies on the OS\_Tree to perform the rank lookups in $O(\log
p)$ time to achieve its running time of $O(p\log p)$. Similarly to the
linked-list version, the first step is to build the data structure. In
this case, it performs $p$ inserts that require $O(\log p)$ time each,
resulting in a $O(p\log p)$ step.


Next is the actual elimination phase. In general, the kid that is
going to be eliminated is $w-1$ ranks above the current caller, where
rank is the order around the circle. So we keep track of the caller at
each iteration, and simple eliminate a kid each time until there is
only one remaining in the tree. After a kid is eliminated the next
caller is going to be the element with rank now equal to what we just
removed (just the next kid in line). So the process repeats, always
making sure not to remove an element with rank that is out of
range. To avoid this, the next caller's rank is always 'modded' with
$1 + $ the remaining kids. This is repeated $p-1$ times so that there
is only kid left in the circle, who is now it. The most expensive
part of this step is finding and removing the correct
element. Searching by rank is $O(\log p)$ given by OS\_Tree, and
remove is also $O(\log p)$ given by the balanced binary tree. Other
operations are in constant time. There are $p-1$ iterations, thus
$p-1$ searches/removes, so the total running time of this
part is $O(p\log p)$.


Both steps of this algorithm run in $O(p\log p)$ time, so the
whole algorithm must be bounded by $p\log p$.

\end{enumerate}

\pagebreak

\section*{Problem 2}

\subsection*{Insert Pseudocode}


\begin{verbatim}

  (* assumed definitions: *)
  (* add_to_end - create p-heap node at end of H with given items and return it *)
  (* get_parent - return parent of given node *)
  (* get_elem - get ith element out of a node *)
  (* swap(node,i,node2,j)  - swap ith element of node with jth element of node2 *)


  (* insert elems (array length p) in to p-heap H *)
  def insert(H, p, elems):
      node = add_to_end(H, elems)
      for i in 1...p:
          bubble_up(node, i, p)

  def bubble_up(node, i, p):
      parent = get_parent(node)
      elem = get_elem(node, i)
      best = infinity
      best_ix = nil
      for j in 1...p:
          parent_elem = get_elem(parent, j)
          if elem > parent_elem and parent_elem < best:
              best = parent_elem
              best_ix = j
      if best_ix is not nil:
          swap(node, i, parent, best_ix)
          bubble_up(parent, best_ix, p)
          
\end{verbatim}

Insert works by creating a new node in the very last position of the
heap that holds all $p$ values to be inserted. Each of these values is
then 'bubbled-up' to the top of the heap just like a normal heap. As
an element is bubbling up, it should only move up if it is greater
than at least one of the values in its parent. The value that it
replaces must be the smallest value in the parent that is smaller than
this element. Because it must check all $p$ value in its parent, each
bubble step is $O(p)$. In the worst case, each inserted value would
need to be bubbled all the way to the root of the tree, which would
mean at most $\log_2 \frac{n}{p}$ bubble ups. So, the worst case
running time for a bubble\_up is $O(p\log_2 \frac{n}{p})$.


A call to insert means $p$ full bubble ups. So the running time of
insert must be $(p^2\log_2 \frac{n}{p})$.

\pagebreak

\subsection*{Extract Max Pseudocode}

\begin{verbatim}

  (* assumed definitions: *)
  (* get_elem - get ith element out of a node *)
  (* swap(node,i,node2,j)  - swap ith element of node with jth element of node2 *)
  (* left/right - get left or right subtree of a node *)
  (* detach_root - remove root node from heap and return it *)
  (* attach_root - set root of the tree to given node *)
  (* detach_last - remove last node from heap and return it *)
  (* items - return items on a node *)

  (* remove p greatest elements in heap *)
  def extract_max(H, p):
      top = detach_root(H)
      last = detach_last(H)
      attach_root(H, last)
      for i in 1...p:
          bubble_down(last, i)
      return items(top)

  def bubble_down(node, i, p):
      best = get_element(node, i)
      best_ix = nil
      best_node = nil
      for j in 1...p:
          check = get_element(left(node), j)
          if check > best:
              best = check
              best_ix = j
              best_node = left(node)
          check = get_element(right(node), j)
          if check > best:
              best = check
              best_ix = j
              best_node = right(node)
      if best_ix is not nil:
          swap(node, i, best_node, best_ix)
          bubble_down(best_node, best_ix, p)

\end{verbatim}

Extract\_max is similar to insert but it has to bubble down the tree
rather than up it. It starts off by removing the root element so its
elements can be returned, and then moves the very last node up and
attaches it to the root. That preserves the 'complete' requirements of
the tree, but the elements on this new root are definitely not the $p$
largest in the tree. So we bubble each $p$ value down the tree to a
spot where it actually fits.


Bubble\_down finds the maximum value out of both it's children's
values first. If this maximum value is greater than the value we are
trying to bubble down, it swaps the two. Now this node satisfies the
heap constraint because we put the largest value in the parent. If we
swapped, there is no guarantee that we didn't screw up the node that
we swapped in to, so we need to try bubbling the value down again. So
we just recur in to the child node at the swapped index. It can only
recurse a maximum of $\log_2 \frac{n}{p}$ times (the height of the
tree). Each bubble\_down call takes $O(p)$ time because we need to
check all $p$ values of both the children nodes to find the
maximum. So, the running time of bubble\_down is $O(p\log_2
\frac{n}{p})$.


The algorithm bubbles down on all $p$ values that are moved to the
root, so we are looking at a running time of
$O(p^2\log_2\frac{n}{p})$.


Both of these algorithms have the same running time if
$O(p^2\log_2\frac{n}{p})$. But $p$ is generally going to be much
smaller than $n$, so we could probably just call it a constant. By
assuming $p$ is constant compared to $n$, the new running time would
be $O(\log_2n)$.

\section*{Problem 3}

The goal of this algorithm is to sort all $m$ strings which have a
total of $n$ characters in $O(n)$ time. To accomplish this, my plan is
use a sort that has aspects of both radix sort and merge sort. 
\newline \newline
If we do a radix sort using the first character as the key, our array
will look something like this:

\begin{center}
  
$\left[
``a...",
``a...",
... ,
``b..." ,
``b..." ,
... ,
``z..."
\right]$
    
\end{center}


\begin{enumerate}[]

\item So the array will now be grouped in to the a's and then the b's
  and then the c's, etc. Each group is in order relative to each
  other. So in the final sorted order, all of the words that start
  with 'a' will still be in those first slots of the array.

\item Now if while we were doing the radix sort, we found the index
  ranges of each 'string-group' at the same time, we wouldn't be
  adding any extra running time to this step. A very convenient
  property of radix sort is that right before we build up our final
  sorted array B, we have the auxillary array C that stores the last
  index of each occurrence in to the final array. After doing the rest
  of the sort, that same array C now stores the first index of each
  occurrence. How does this help?  Well those values are the ranges of
  what we were trying to keep track of!

\item In C's first state, C['a'] is the index of the last occurrence
  of a string starting with 'a' in the final sorted array. Likewise
  for all other characters. After the sort is finished, C['a'] is now
  the first occurrence of a string starting with 'a' in the final
  sorted array. We just easily found the indices of all strings
  starting with 'a' in our final sorted array.

\item We can now recursively call our sorting algorithm on each subset
  of the array that we just found. If we sort all of the a's only with
  respect to each other, all the b's with respect to each other,
  etc. we will have a completely sorted array.

\item Each level of recursion will sort on the next character position
  unless a group has only one member, because an array of size one is
  sorted trivially. 

\item Each time a group is formed on a nul character that was used for
  padding, we don't need to sort that group because it is in it's
  final position in the array. So on the next level of recursion, we
  don't have to keep comparing against a nul character which would end
  up making the running time worse than $O(n)$.

\item Some rough pseudocode should solidify the algorithm.
  
  \begin{verbatim}

    (* g_radix_sort is the basic radix sort that returns the
    string-groups described above. The format of the string
    groups would be an array whose elements are the indices of the
    first and last occurrence of a string starting with that
    character position. The last group (27th index) represents
    the group starting with the nul character *)

    (* sort arr from index first to last using radix style sort on key *)
    def str_sort(arr, first, last, key):
        groups = g_radix_sort(arr, first, last, key)
        for i in 1...26:
            front = groups[i][1]
            back = groups[i][2]
            if (back - front) > 1:
                str_sort(arr, front, back, key+1)
            
  \end{verbatim}

\item Worst case analysis: The worst case of this sort would be if all
  m strings were identical up until the very last character. In this
  scenario, each string would be of length $\frac{n}{m}$. If we let
  that length be represented by the $l$, then we know for the first
  $l-1$ calls to this function, we aren't making the problem any
  smaller. 

  
\end{enumerate}



\end{document}
